
%% Introduction page
%% =============
%%
Volcanic activities on Earth have always shaped the earth surface and influenced atmospheric processes. Volcanoes are often particularly recognized by their dramatic consequences of a major volcanic eruption. But volcanoes influence our lives in more than this way. Volcanic gases can effect the weather \textcolor{green}{by emitting aerosols}(timescales of days to weeks) or the climate (timescales of months to years) \cite{schmidt2015volcanismarticle}.
Examples are the Laki eruption in Iceland (1783-1784) followed by a very hot summer and a cold winter in central Europa \cite{thordarson2003atmospheric} and the Tambora eruption, indonesia in 1815 which caused the "year without summer" in 1816.\\
%
\newline
%
Considering plate tectonics of earth  most volcanoes are caused by diverging or converging of the continental plates and therefore located at the margins of the continental plates.\textcolor{blue}{as well as during passing degassing phases}
\textcolor{red}{Ich wuerde das so nicht schreiben - ich meine die plattentektonik ist ja auch nur der Waeremkonvektion und gravitaion im innern der Erde geschulded nd so eher indirekte (zweitrangige ursache des Vulcanismus. Zu schreiben das Vulkane an Plattenraendern zu finden sind ja - natuerlich }
Another possibility for occurrence of volcanoes is the the interior of continental or oceanic shelves. \cite{schmincke2000vulkanismus}\\
The most abundant volatile species released during a volcanic eruption are water vapour (H$_2$O; relative amount of the plume: 50\%-90\%) and carbon dioxide (CO$_2$; relative amount of the plume: 1\%-40\%) \cite{platt2015quantification}. But the short effects of those two gases are rather small since their effect on atmospheric composition is negligibly due to the high abundance of atmospheric H$_2$O and CO$_2$. But on timescales of the age of the earth the volcanic emission of H$_2$O and CO$_2$ are the source of our current atmosphere. \cite{schmidt2015volcanism}\\ 
\textcolor{red}{CO2-Emissionen sind schon auf viel kürzeren Zeitskalen (einige 100000 Jahre) relevant für die Erhaltung des Kohlenstoffs (CO2, Biomasse) an der Erdoberfläche.}
\textcolor{red}{auch hier bitte etwas vorsichtiger - du kannst gern schriebn das Vulkane largly contributed on the formation of our Atmosphere	aber so wie es jetzt da steht stehen sie als einzige quellen da und das ist so auch wieder nicht ganz richtig - das wuerden die astrophysiker in Heidelberg nicht gern hoeren}
A typically volcanic plume consists of many different gases alongside H$_2$O and CO$_2$  sulfur dioxide (SO$_2$) contributes with 1\%-25\% to the plume, hydrogen sulfide (H$_2$S) with 1\%-10\% and hydrogen chloride with (HCl) 1\%-10\%. Furthermore there are trace gases for example carbon disulfide (CS$_2$), carbon sulfide (COS) carbon monoxide (CO) hydrogen fluoride (HF), hydrogen bromide (HBr) and many other species. \cite{platt2015quantification}\\
%
A decrease of stratospheric ozone (O$_3$) has been observed after the eruption of  El Chickon in 1982 and the eruption of Mount Pinatubo 1991. A depletion stratospheric O$_3$ results in ozone holes.\textcolor{red}{kann auch wider zu missverstaendissen fuehren - Ozone hole ist ein definierter begriff nicht jede O3 abnahme ist gleich ein Ozonloch} The depletion comes from volcanic aerosols which serve to transform anthropogenic chlorine/bromine into more reactive forms \cite{solomon1998ozone}. 
%
Volcanic gases can alter the radiative balance of the earth in timescales relevant for climate change \textcolor{blue}{??? vertsehe ich nicht was sind denn zeitskalen relevant fuer Klimaaenderung} due to scatter and absorption of solar radiation \cite{schmidt2015volcanism}.\\
%
The gas composition of the volcano plume change with activity and can be a indication for the processes inside the earth.\\ 
%
In this work we are particularly interested in the ratio of BrO and SO$_2$.
\textcolor{red}{hier gehts jetzt zu schnell - das muss Du fuer den unerfahrenene leser genauer erklaeren} The halogen sulfur ratio is a proxy for volcanic processes. Therefore we make the assumption that the ratio of BrO and \ce{SO2} contains informations about its degassing source depth.\textcolor{red}{Das ist doch keine "Annahme", es gibt gute Gründe für diesen Zusammenhang.} A change in BrO/\ce{SO2} prior to eruption was observed at Etna and Nevado del Ruiz.\\
%
\newline
%
The data of this thesis are recorded by the Network for Observation of Volcanic and Atmospheric Change (NOVAC) which is a network to record volcanic emissions. The aim of NOVAC is mainly to provide new parameters for risk assessment and volcanological research, both locally and on a regional and global scale.
NOVAC is a Network of spectrographic instruments located next to about 30 volcanoes in Asia, America, Africa and Europe. At each of these volcanoes there are two to four spectrograph's installed, recording back-scattered solar radiation spectra at different viewing angles.\\
NOVAC is a network which produces a large amount of data and we have the chance to evaluate long time periods which is a unique opportunity to study volcanic trace gases.\textcolor{green}{muss das noch genauer?}\\
Since the conditions at volcanoes are rough, the instruments need to be rather simple to keep the maintenance cheap and to assure a longer lifetime of the instruments. \textcolor{green}{Hier geht die Argumentation etwas durcheinander: Die Instrumente selbst und ihre Wartung müssen preiswert sein, daher einfach.} Therefore it was decided not to implement temperature stabilization even at the expense of the quality of the data.\\
%
\newline
%
The data recorded by NOVAC are evaluated using Differential Optical Absorption Spectroscopy (DOAS) \cite{platt2008differential}. DOAS exploit the wavelength dependency of the absorption of light and is based on the lambert-beer law. The gas concentration is retrieved from the characteristic difference in absorption structures between the plume and a reference region. Thus it is fundamental to have a reference which is free of the gases of interest, to assure that the retrieved concentration is correct.
%
\newline
%
The reference region, is usually treated as free of
volcanic trace gases. If the reference region is for any reason
contaminated by volcanic trace gases, the reference spectrum has to be
replaced by a volcanic-gas-free reference. Alternative spectra could be for example a
theoretical solar atlas spectrum or a volcanic-gas-free reference
spectrum recorded in the temporal proximity (eg. a day before) by the same instrument.
The first option comes with the drawback of reduced precision, as the
instrumental effects have to be modeled and added to the retrieval. The reduction in precision is acceptable for the \ce{SO2} retrieval, but not suitable for a BrO retrieval because then most data would be below the detection limit. The advantage of the first option is, that the absolute column density is calculated, since, the theoretical solar atlas spectrum is absolutely gas free. For the second option, the alternative reference spectrum should
have been recorded at similar conditions with respect to meteorology,
radiation,intensity and in temporal proximity due to instrumental changes
with time and ambient conditions. We combined both options in order to
achieve both, enhanced accuracy but still maximum possible precision of
the \ce{SO2} and BrO retrievals. We present an algorithm which finds the
optimal reference spectrum automatically. As first step, a possible \ce{SO2}
contamination of the standard reference is checked by a comparison with
the theoretical solar atlas. If a contamination is detected, as second step,
the algorithm picks a volcanic-gas-free reference (beforehand
automatically checked for contamination) from another scan.\\
%
\newline
%
In this work we are mainly dealing with data from Tungurahua in Ecuador and Nevado del Ruiz a volcano located in Colombia.